---
title: "Report analysis for niche of sister species pairs and clades"
output: word_document
---
###################################################
### Clade niche analysis
###################################################


```{r libraries echo=FALSE}
library(phytools)
library(ape)

```

##############################################################################
### Data preparation
##############################################################################

```{r echo=FALSE}
setwd("C://Users//nomur//Documents")
source(".//Acaena niche evolution//06_Clade pairing.R")
########################################################################################
### Phylogenetic distances between nodes
########################################################################################
library(dplyr)
library(phytools)
library(nichePlot)

# Import phylogeny tree data
acaena <- read.nexus("Y:\\Niche change of lineages\\Niche evolution of open habitat species in islands\\Phylogenetic data and trees\\From Angela\\NZ_Acaena_BEAST_output_6gene.tree")

########################################################################################
### Get sister nodes
########################################################################################

### Get sister species (tip) or sister group of a target species
# Extract names of edges (i.e. tips and taxa)
tips <- acaena$tip.label

## first get the node numbers of the tips
nodes <- data.frame(sapply(tips, function(x,y) which(y==x), y=acaena$tip.label))
colnames(nodes)<-"nodelabel"

# Modify species names in phylogentic distance file
a <- unlist(strsplit(rownames(nodes), "_Ac"))
a2 <- gsub("_EU352216", "", a) %>% gsub("_AY634821", "", .) %>% gsub("novae-", "novae.", .)
rownames(nodes) <- grepl("_", a2) %>% a2[.]

## Second, find sister node of a target species

tipssister <- findSisterNode(acaena)

# Name a result list
names(tipssister) <- rownames(nodes)

### Get a sister group of internal nodes as well as terminal nodes

# Get node numbers of internal nodes
allnodesister <- GetInternalNodeNumber(acaena)

allnodesister[1:length(tipssister)] <- tipssister

########################################################################################
### Plot edge length on trees
########################################################################################

### Phylogenetic distance between nodes
# The order of the node pair doesn't matter. dist.nodes(acaena)[i, getSisters(acaena, i)] == dist.nodes(acaena)[getSisters(acaena, i), i]
distance <- sapply(1:max(acaena$edge), function(i){
  c(i, getSisters(acaena, i), dist.nodes(acaena)[i, getSisters(acaena, i)])
})

distance2 <- do.call(rbind, distance)
colnames(distance2) <- c("node", "sisterNode","distance")

```

#################################################################################
### Calculate node ages
#################################################################################
```{r echo=FALSE}
### Distance between all combinations of tips
distances <- dist.nodes(acaena)

# Calculate species ages
ages <- sapply(nodes[[1]], function(i){
  # Phylogenetic distance list has 0 (distance to themselves)
  (distances[,i] > 0) %>% distances[., i] %>% min
  }
)

ages <- as_tibble(ages) %>% mutate(., spname = row.names(nodes))

### Calculate internal node ages
nodeage <- cbind(rep(NA, length(branching.times(acaena))), branching.times(acaena))
colnames(nodeage) <- c("spname", "value")


agesTipAndNode <- rbind(ages, nodeage)


```


########################################
###  Dataframe of Clade niche volume & age
########################################
```{r echo=FALSE}
volume <- read.csv(".//clade_nicheVolume_acaena.csv")
extractAges <- agesTipAndNode[rownames(agesTipAndNode) %in% volume$node1,]

ageVolData <- cbind(extractAges, volume[order(volume$node1), ])

```



###################################################
### Dataframe of Clade niche overlap & phylogenetic distances
###################################################
```{r echo=FALSE}
dis <- data.frame(distance2)
overlap <- read.csv(".//clade_schoennerD_acaena.csv")

overlapPdData <- cbind(overlap, dis[dis$node %in% overlap$node1, ])[ ,c("node1", "node2", 
                                                      "ecospat.corrected", "ecospat.uncorrected", "distance")]
```


#########################################################################
### Plots
#########################################################################
```{r}
########################################################
### Function for plotting
########################################################

# This fucntion was extracted from "10_Analysis clade niche.R" for Acaena sister clades

library(ggplot2)

plotAnalysis <- function(data, 
                         m, # linear model object
                         xv, yv, # column names of responce and explanatory variable
                         xlabname, ylabname, # axes names for plot
                         nodeNumber,
                         showStats = T # TRUE; Show p value and slope of linear model and colour points, FALSE; No stat values and black points
){
  
  if(showStats == T){
    myplot <- ggplot(data, aes_string(x = xv, y = yv, label = nodeNumber)) +
      geom_point() +
      # text label for points
      geom_text(size=5) +
      # change xy labels
      labs(x = xlabname, y = ylabname) +
      # change text size
      theme(text = element_text(size = 20),
            axis.text.x = element_text(size = 20)) +
      # drow LM line & confident intervals 
      stat_smooth(method = "lm", col = "red") +
      # show stats result as title
      labs(title = paste("Adj R2 =", signif(summary(m)$adj.r.squared, digits = 2),
                         "Intercept =", signif(m$coef[[1]], 2),
                         " Slope =", signif(m$coef[[2]], 2),
                         " P =", signif(summary(m)$coef[2, 4], 2))) +
      theme(panel.background = element_rect(fill = "gray95"))
  } else {
    myplot <- ggplot(data, aes_string(x = xv, y = yv)) +
      geom_point() +
      # change xy labels
      labs(x = xlabname, y = ylabname) +
      # change text size
      theme(text = element_text(size = 20),
            axis.text.x = element_text(size = 20)) +
      # drow LM line & confident intervals 
      stat_smooth(method = "lm", col = "red") +
      theme(panel.background = element_rect(fill = "gray95"), legend.position="none")
  }
  return(myplot)
}

```

#########################################################################
### Sister species pairs' Phylogenetic distances ~ niche overlap of occurrence records
#########################################################################

```{r echo=FALSE}
### Node numbers of sister species pairs
sispairs <- c(1,20,3,5,16)
sisOverlapPd <- (overlapPdData$node1 %in% sispairs) %>% overlapPdData[., ]

m <- lm(distance ~ ecospat.corrected, sisOverlapPd)

myplot <- plotAnalysis(data = sisOverlapPd, 
                       m = m, 
                       xv = "ecospat.corrected", yv = "distance", 
                       nodeNumber = "node1", showStats = T,
                       xlabname = "Niche overlap of occurrence records", ylabname = "Phylogenetic distances between sister species pairs"
)

plot(myplot)
# save
# ggsave(paste("Y:\\sister_pd_nicheoverlap_legend.png", sep = ""), plot = myplot,
#        width = 300, height = 210, units = 'mm')

rm(myplot, m)

```

#########################################################################
### Sister species pair - their ancestor's sister node
### Phylogenetic distances ~ niche overlap of occurrence records
#########################################################################
```{r}

### Node numbers of sister species pairs
sispairs <- c(1,20,3,5,16)

ancSisNode <- sapply(sispairs, function(i){
  
  ancestor <- acaena$edge[which(i == acaena$edge[, 2])]
  ancestorSisNode <- distance2[distance2[,"node"] == ancestor, "sisterNode"]
  return(ancestorSisNode)
  
  }
)


ancsisOverlapPd <- rbind(
  ((overlapPdData$node1 %in% ancSisNode) %>% overlapPdData[., ]),
  ((overlapPdData$node2 %in% ancSisNode) %>% overlapPdData[., ])
)

dup <- duplicated(ancsisOverlapPd$ecospat.corrected) %>% which
ancsisOverlapPd <- ancsisOverlapPd[-dup, ]

m <- lm(distance ~ ecospat.corrected, ancsisOverlapPd)

myplot <- plotAnalysis(data =ancsisOverlapPd, 
                       m = m, 
                       xv = "ecospat.corrected", yv = "distance", 
                       nodeNumber = "node1", showStats = T,
                       xlabname = "Niche overlap of occurrence records", ylabname = "Phylogenetic distances between sister species pairs"
)

plot(myplot)

# # save
# ggsave(paste("Y:\\sisterAunt_pd_nicheoverlap_legend.png", sep = ""), plot = myplot,
#        width = 300, height = 210, units = 'mm')

rm(myplot, m)

```

#########################################################################
### Clade Phylogenetic distances ~ niche overlap of occurrence records
#########################################################################
```{r}

### Eliminate outlier, SAC, of clade age

overlapPd <- overlapPdData[which(overlapPdData$node1 != 19), ]
m <- lm(distance ~ ecospat.corrected, overlapPd)

myplot <- plotAnalysis(data = overlapPd, 
                       m = m, 
                       xv = "ecospat.corrected", yv = "distance", 
                       nodeNumber = "node1", showStats = T,
                       xlabname = "Niche overlap of occurrence records", ylabname = "Phylogenetic distances between clades"
                       )
plot(myplot)

# # save
# ggsave(paste("Y:\\clade_pd_nicheoverlap_legend.png", sep = ""), plot = myplot,
#        width = 300, height = 210, units = 'mm')
# 
# rm(myplot, m)
# 
# 
# ### Leave outlier of clade age in data
# m <- lm(distance ~ ecospat.corrected, overlapPdData)
# 
# myplot <- plotAnalysis(data = overlapPdData, m = m, xv = "ecospat.corrected", yv = "distance",
#                        nodeNumber = "node1", showStats = T,
#                        xlabname = "Niche overlap of occurrence records", ylabname = "Phylogenetic distances between clades")
# 
# plot(myplot)
# 
# # # save
# # ggsave(paste("Y:\\clade_pd_nicheoverlap_outlier.png", sep = ""), plot = myplot,
# #        width = 300, height = 210, units = 'mm')
# 
# rm(myplot, m)

```

#########################################################################
### Clade ages ~ niche volume of occurrence records
#########################################################################

```{r}

### Eliminate outlier
outlier <- which(max(ageVolData$value) == ageVolData$value)
ageVol <- ageVolData[-outlier,]

m <- lm(value ~ ecospat.corrected, ageVol)

myplot <- plotAnalysis(data=ageVolData[-outlier,], m=m, xv = "ecospat.corrected", yv = "value", 
                       nodeNumber = "node1", showStats = T,
                       xlabname = "Niche volume of occurrence records", ylabname = "Clade age")

plot(myplot)

# # save
# ggsave(paste("Y:\\clade_age_nicheVolume_legend.png", sep = ""), plot = myplot,
#        width = 300, height = 210, units = 'mm')
# 
# rm(myplot, m)
# 
# 
# ### Leave outlier
# m <- lm(value ~ ecospat.corrected, ageVolData)
# 
# myplot <- plotAnalysis(data=ageVolData, m=m, xv = "ecospat.corrected", yv = "value", 
#                        nodeNumber = "node1", showStats = T,
#                        xlabname = "Niche volume of occurrence records", ylabname = "Clade age")
# 
# # save
# ggsave(paste("Y:\\clade_age_nicheVolume_outlier.png", sep = ""), plot = myplot,
#        width = 300, height = 210, units = 'mm')
# 
# rm(myplot, m)

```


